---
title: "Collection and Cleaning 2018"
author: "Jake Harmon"
date: "1/25/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidylog)
library(rvest) #for the html_session
library(wdman) #selenium binary manager
library(RSelenium)
library(xml2)
library(selectr)
```

SPOTRAC Authentication
```{r}
loginpage <- "https://www.spotrac.com/signin/?redirect="
pgsession <- html_session(loginpage)
pgform <- html_form(pgsession)[[3]] 
filled_form <- set_values(pgform,
                          email="jacobwharmon@gmail.com", 
                          password="qy3vZLM9mD!RQTx")
submit_form(pgsession, filled_form)
```

# CAP Team Info by Year: https://www.spotrac.com/nba/<team>/cap/<year>     2020 means current
# CAP League Info by Year: https://www.spotrac.com/nba/cap/<year> 
# CASH Player All time https://www.spotrac.com/nba/<team>/<player_id>/cash-earnings/

Grabbing all player URL's
```{r}
teams_SPO <- c("atlanta-hawks","boston-celtics","brooklyn-nets","charlotte-hornets","chicago-bulls",
              "cleveland-cavaliers","dallas-mavericks","denver-nuggets","detroit-pistons","golden-state-warriors",
              "houston-rockets","indiana-pacers","los-angeles-clippers","los-angeles-lakers","memphis-grizzlies",
              "miami-heat","milwaukee-bucks","minnesota-timberwolves","new-orleans-pelicans","new-york-knicks",
              "oklahoma-city-thunder","orlando-magic","philadelphia-76ers","phoenix-suns","portland-trail-blazers",
              "sacramento-kings","san-antonio-spurs","toronto-raptors","utah-jazz","washington-wizards")

years_SPO <- c(2018:2020) #only works for 2020,2019,2018. The HTML is different for the 2017 and older years

player_url_list <- list()

for (i in c(1:30)) { #This works for 2020 and 2019
  temp_team <- teams_SPO[i]
  player_url_list[[paste0(temp_team,years_SPO[1])]] <- read_html(paste0("https://www.spotrac.com/nba/",
                                                                       temp_team,
                                                                       "/cap/",
                                                                       years_SPO[1])) %>%
                                                                       querySelector("div.team-content") %>%
                                                                       querySelector("div.teams") %>%
                                                                       html_nodes("table") %>%
                                                                       html_nodes("tr") %>%
                                                                       html_nodes("td") %>%
                                                                       html_nodes("a") %>%
                                                                       html_attr("href")
  print(paste0(i, "out of 30 completed"))
}
remove(temp_team)
```

#Note, each team can only have 15 players even though more player links got grabbed
Grabbing Player Names, Stats
# https://www.spotrac.com/nba/<team>/<player_id>/statistics/
```{r}
initial_url <- paste0(player_url_list[[paste0(teams_SPO[1],years_SPO[1])]][1],"statistics/")

initial_team <- teams_SPO[1]

initial_year <- years_SPO[1]

initial_name <- read_html(initial_url) %>% 
              querySelector("div.player-details") %>% 
              querySelector("div.player-name") %>% 
              html_nodes("h1") %>% 
              map_chr(xml_text) %>% 
              str_trim("both")

initial_id <- as.numeric(str_remove(str_remove(player_url_list[[paste0(teams_SPO[1],years_SPO[1])]][1],
                                               "https://www.spotrac.com/redirect/player/"),"/"))

player_names_ids <- data.frame("team" = initial_team,"year" = initial_year,
                               "name" = initial_name,"id" = initial_id)
stat_tables <- list()

for (i in c(25:30)){ #Breaks at OKC (team 21/30) 4/15 at Schroder and PHX (team 24/30) 12/15 at Elie Okobo
  print(teams_SPO[i])
  for (k in c(1:15)){
    temp_year <- years_SPO[1]
    temp_team <- teams_SPO[i]
    temp_url <- paste0(player_url_list[[paste0(temp_team,temp_year)]][k], #k for 1:15 players
                       "statistics/")
    
    temp_name <- read_html(temp_url) %>% 
                  querySelector("div.player-details") %>% 
                  querySelector("div.player-name") %>% 
                  html_nodes("h1") %>% 
                  map_chr(xml_text) %>% 
                  str_trim("both")
    
    temp_id <- as.numeric(str_remove(str_remove(player_url_list[[paste0(temp_team,temp_year)]][k],
                                                   "https://www.spotrac.com/redirect/player/"),"/"))
    
    player_names_ids <- rbind(player_names_ids, c("team" = temp_team, "year" = temp_year,
                                                  "name" = temp_name, "id" = temp_id))
    
    try(stat_tables[[as.character(temp_id)]] <- read_html(temp_url) %>% 
                            querySelector("li#statistics") %>% 
                            querySelector("table.salaryTable") %>% 
                            html_nodes("tr") %>% 
                            html_nodes("td") %>% 
                            map_chr(xml_text), silent = TRUE) #16 columns that need to get wrapped
    print(paste(k,"out of 15"))
  }
}

remove(i,k,temp_url,temp_name,temp_id,temp_team,temp_year,
       initial_url,initial_name,initial_id,initial_team,initial_year)

player_names_ids <- player_names_ids %>% unique()  
```


Unwrapping the vectorized stat_tables entries and creating giant playerID-year stat table
```{r}
init_stat_df <- data.frame(t(stat_tables[[1]][1:16]))
names(init_stat_df) <- c("season","team","games","mpg","fg%","3pt%","ft%","orb","drb",
                         "trb","ast","stl","blk","tov","pt","prod_pts")

unwrap_fx <- function(x){
  numrow <- length(x) / 16
  temp_df <- data.frame(matrix(NA, nrow = 0, ncol = 16))
  names(temp_df) <- c("season","team","games","mpg","fg%","3pt%","ft%","orb",
                      "drb","trb","ast","stl","blk","tov","pt","prod_pts")
  for(k in c(1:numrow)){
    which_rows <- c(1:16)
    temp_df[k, ] <- x[which_rows + (16*(k-1))]
  }
  return(temp_df)
}

# for each index across stat_tables, create a column equal to the playerID and then add their stat table

big_stat_table <- data.frame(matrix(NA, nrow = 0, ncol = 17))
names(big_stat_table) <- c("player_id","season","team","games","mpg","fg%","3pt%","ft%","orb",
                                    "drb","trb","ast","stl","blk","tov","pt","prod_pts")
player_id_vec <- names(stat_tables)

for (i in c(1:442)) {
  temp_df <- unwrap_fx(stat_tables[[i]])
  temp_df$player_id <- player_id_vec[i]
  temp_df <- temp_df %>% select(player_id, everything())
  big_stat_table <- rbind(big_stat_table, temp_df)
}

big_stat_table <- big_stat_table %>% filter(season != "Totals")

big_stat_table <-left_join(big_stat_table, player_names_ids,
                           by = c("player_id" = "id")) %>% select("name",everything(), -year, -team.y)

write_csv(big_stat_table, "big_stat_table_2018.csv")
write_csv(player_names_ids, "player_names_ids_2018.csv")
```

Dennis Schröder
```{r}
    temp_year <- years_SPO[1]
    temp_team <- teams_SPO[21]
    temp_url <- paste0(player_url_list[[paste0(temp_team,temp_year)]][4], #k for 1:15 players
                       "statistics/")
    
    temp_name <- "Dennis Schröder"
    
    temp_id <- as.numeric(str_remove(str_remove(player_url_list[[paste0(temp_team,temp_year)]][4],
                                                   "https://www.spotrac.com/redirect/player/"),"/"))
    
    player_names_ids <- rbind(player_names_ids, c("team" = temp_team, "year" = temp_year,
                                                  "name" = temp_name, "id" = temp_id))
    
    try(stat_tables[[as.character(temp_id)]] <- read_html(temp_url) %>% 
                            querySelector("li#statistics") %>% 
                            querySelector("table.salaryTable") %>% 
                            html_nodes("tr") %>% 
                            html_nodes("td") %>% 
                            map_chr(xml_text), silent = TRUE)
```

Élie Okobo
```{r}
    temp_year <- years_SPO[1]
    temp_team <- teams_SPO[24]
    temp_url <- paste0(player_url_list[[paste0(temp_team,temp_year)]][12], #k for 1:15 players
                       "statistics/")
    
    temp_name <- "Élie Okobo"
    
    temp_id <- as.numeric(str_remove(str_remove(player_url_list[[paste0(temp_team,temp_year)]][13],
                                                   "https://www.spotrac.com/redirect/player/"),"/"))
    
    player_names_ids <- rbind(player_names_ids, c("team" = temp_team, "year" = temp_year,
                                                  "name" = temp_name, "id" = temp_id))
    
    try(stat_tables[[as.character(temp_id)]] <- read_html(temp_url) %>% 
                            querySelector("li#statistics") %>% 
                            querySelector("table.salaryTable") %>% 
                            html_nodes("tr") %>% 
                            html_nodes("td") %>% 
                            map_chr(xml_text), silent = TRUE)
```


Grabbing Player Contract Info: VIP
```{r}
temp_contract_table <- read_html(temp_url) %>% 
                        querySelector("li#contracts") %>% 
                        querySelector("table.salaryTable") %>% 
                        html_nodes("tr") %>% 
                        html_nodes("td") %>% 
                        map_chr(xml_text) #grabs the current contract info
```



Error Checking
```{r}

```



















