---
title: "Building the March Madness Dataframe"
author: "Ken Harmon"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:  
    keep_md: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
editor_options: 
  chunk_output_type: console
---

# {.tabset .tabset-fade}

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
pacman::p_load(tidyverse, DT)
theme_set(theme_bw())
```

```{r swd, eval=FALSE, echo=FALSE}
# this is set to not run during the knit process
# this sets the working directory to the file location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```



## Games 85 to 19

Here are all the march Madness games that have been played since 1985 when March madness first went to a 64 team bracket.

This includes play in games.

```{r atg}
games85 <- read.csv("All_NCAA_Tournament_Games85to19.csv") %>% select(c(2,3,5,6,8:10,12))

datatable(games85, rownames = FALSE ,extensions = "Responsive",options=list(lengthMenu = c(10,25,68)))
```

## Rankings and stats

NCAA Season Stats: https://www.sports-reference.com/cbb/seasons/

All Tournament Game Results:

  From which I can get Seed Expectations
  From Which I can get each teams PASE
  
Kenpom Rankings: https://kenpom.com/index.php?s=TeamName

Here are the team season stats since 2000

```{r ras}
teamstats00 <- read.csv("NCAASeasonTeamStats00to19.csv", na.strings=c("","NA")) 

datatable(teamstats00, rownames = FALSE ,extensions = "Responsive",options=list(lengthMenu = c(10,25,68)))
```

## PASE

Here are the all-time 

Performance Against Seed Expectations (PASE)

and

Performance Against KenPom Expectations (PAKE)

```{r pase}
PASE <- read.csv("AllTimePAKE.csv", na.strings=c("","NA")) %>% select(1:3)

datatable(PASE, rownames = FALSE ,extensions = "Responsive",options=list(lengthMenu = c(10,25,68)))
```

## KenPom

Here are the KenPom Rankings since 2002

```{r kp}
KP02 <- read.csv("KenPomNCAARankings02to19.csv", na.strings=c("","NA")) %>% select(1,3,4,5) 

datatable(KP02, rownames = FALSE ,extensions = "Responsive",options=list(lengthMenu = c(10,25,68)))
```


## Favored

Brake the data into Favored and Underdog

```{r fvu}
all02 <- left_join(games85,KP02, by = c("Winner"="Team", "Year"="Year"))%>%filter(Year>2001)
all02 <- left_join(all02,KP02, by = c("Loser"="Team", "Year"="Year"))
write.csv(all02, "gameswkp.csv", row.names = FALSE)
```

## Reload

Here are all the games by Favored and Underdog with stats

```{r rg}
FUstats <- read.csv("FUGames02.csv", na.strings=c("","NA"))

FUstats <- left_join(FUstats,PASE, by = c("Favored"="TEAM"))
FUstats <- left_join(FUstats,teamstats00, by = c("Favored"="Team", "Year"="Year")) 
FUstats <- left_join(FUstats,PASE, by = c("Underdog"="TEAM"))

FUstats <- left_join(FUstats,teamstats00, by = c("Underdog"="Team", "Year"="Year")) %>% select(1:11,14:46,49:79)
#colnames(all02)[9:43] <- paste("w.", colnames(all02[,c(9:43)]), sep = "")
datatable(FUstats, rownames = FALSE ,extensions = "Responsive",options=list(lengthMenu = c(10,25,68)))
```

## Final Out

Here is the final Output Set

```{r export}
#write.csv(FUstats, "Prediff.csv", row.names = FALSE)

```



