---
title: "Another PCA"
author: "Ken Harmon"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:  
    keep_md: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
editor_options: 
  chunk_output_type: console
---

# {.tabset .tabset-fade}

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
#install.packages("tidyverse") # run this line once in console to get package

library(tidyverse)
theme_set(theme_bw())
```

## The Dataset

https://www.analyticsvidhya.com/blog/2016/03/practical-guide-principal-component-analysis-python/

```{r data}
#load train and test file
train <- read.csv("TrainBMS.csv")
test <- read.csv("TestBMS.csv")

#add a column
test$Item_Outlet_Sales <- 1

#combine the data set
combi <- rbind(train, test)

#impute missing values (item weight) with median weight
combi$Item_Weight[is.na(combi$Item_Weight)] <- median(combi$Item_Weight, na.rm = TRUE)

#impute 0 with median
combi$Item_Visibility <- ifelse(combi$Item_Visibility == 0, median(combi$Item_Visibility),combi$Item_Visibility)

#find mode and impute
table(combi$Outlet_Size, combi$Outlet_Type)
levels(combi$Outlet_Size)[1] <- "Other"
```

Till here, we’ve imputed missing values. Now we are left with removing the dependent (response) variable and other identifier variables( if any). As we said above, we are practicing an unsupervised learning technique, hence response variable must be removed.

```{r slim}
#remove the dependent and identifier variables
my_data <- subset(combi, select = -c(Item_Identifier, Outlet_Identifier, Item_Outlet_Sales))
```

Sadly, 6 out of 9 variables are categorical in nature. We have some additional work to do now. We’ll convert these categorical variables into numeric using one hot encoding.

```{r Categoricaltonum}
library(dummies)
#create a dummy data frame
new_my_data <- dummy.data.frame(my_data, names = c("Item_Fat_Content","Item_Type","Outlet_Establishment_Year","Outlet_Size","Outlet_Location_Type","Outlet_Type"))
```

```{r tt}
#divide the new data
pca.train <- new_my_data[1:nrow(train),]
pca.test <- new_my_data[-(1:nrow(train)),]
```

```{r pca}
#principal component analysis
prin_comp <- prcomp(pca.train, scale. = T)
names(prin_comp)
#outputs the mean of variables
prin_comp$center

#outputs the standard deviation of variables
prin_comp$scale

prin_comp$rotation

prin_comp$rotation[1:5,1:4]

dim(prin_comp$x)

biplot(prin_comp, scale = 0)

```

```{r}
train.data <- data.frame(Item_Outlet_Sales = train$Item_Outlet_Sales, prin_comp$x)
train.data <- train.data[,1:31]

library(rpart)
rpart.model <- rpart(Item_Outlet_Sales ~ .,data = train.data, method = "anova")
rpart.model

test.data <- predict(prin_comp, newdata = pca.test)
test.data <- as.data.frame(test.data)
test.data <- test.data[,1:30]

rpart.prediction <- predict(rpart.model, test.data)
```


