---
title: "SIGACTS Western Yemen"
author: "Ken Harmon"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:  
    keep_md: true
    code_folding: hide
    fig_height: 6
    fig_width: 12
    fig_align: 'center'
editor_options: 
  chunk_output_type: console
---

# {.tabset .tabset-fade}

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,fig.width=12, fig.height=8)
```


maptools: a set of tools for reading and handling spatial objects. In particular, it will be used to read .shp files, a common format used by geographic information systems software.
ggplot2: one of the powerful graphics engines available to R users
ggmap: a package for spatial visualization using popular on-line mapping systems, such as GoogleMaps or OpenStreetMap.


```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
# devtools::install_gitlab("hrbrmstr/mgrs")

pacman::p_load(tidyverse, ggmap, sp, tmap, maps, maptools, raster, rgdal, PBSmapping)
theme_set(theme_bw())
register_google(key = "AIzaSyAfnLNZjvYdMx-cyga_qA1oJ6P36dRGalA") 
```

```{r swd, eval=FALSE, echo=FALSE}
# this is set to not run during the knit process
# this sets the working directory to the file location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```




```{r lgs}
load(file = "sigacts.rda")
load(file = "wy_map.rda")
load(file = "countries_spdf.rda")
load(file = "countries_sp.rda")

# Use base_layer argument to ggmap() and add facet_wrap()
  ggmap(WY_8, 
    base_layer = ggplot(sigacts %>% filter(attacktype1 == 3), aes(lon, lat))) +
    #geom_polygon(data = countries_spdf_df, aes(x = long, y = lat, group = group, alpha = .1, color = "id")) +
  geom_point(aes(color = as.factor(iyear), size = nkill)) #+
  #scale_x_continuous(limits = c(43, 44), expand = c(0, 0)) #+
  #scale_y_continuous(limits = c(10, 20), expand = c(0, 0)) +
  #facet_wrap(vars(attacktype1_txt))
    
```

```{r wcm}
library(broom)
countries_spdf_df <- tidy(countries_spdf)

countries_spdf$name <- sapply(slot(countries_spdf, "polygons"), function(x) slot(x, "ID"))

countries_spdf_df <- merge(countries_spdf_df, countries_spdf, by.x = "id", by.y = "name")

ggmap(wy_map) +
  geom_polygon(data = countries_spdf_df, aes(x = long, y = lat, group = group, alpha = .1 ))

```

