---
title: "Earth as a Perfect Spheroid"
author: "Ken Harmon"
date: "`r format(Sys.time(), '%Y %B %d')`"
output:
  html_document:
    code_folding: hide
    fig_align: center
    fig_height: 6
    fig_width: 12
    keep_md: yes
editor_options:
  chunk_output_type: console
---

# {.tabset .tabset-fade}

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load_libraries, include=FALSE}
# Use this R-Chunk to load all your libraries!
pacman::p_load(tidyverse, mgrs, DT, numbers,numform)

theme_set(theme_bw())
```

```{r swd, eval=FALSE, echo=FALSE}
# this is set to not run during the knit process
# this sets the working directory to the file location
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

### What if I assume the earth is a perfent spheroid?

## Premise

Then I should not be able to convert Longitude and Latitude coordinates to MGRS with any type of accuracy. The geologic science people tell me this cannot be done.

Challenge accepted.

## Dataset

Here is my initial dataset of random points around the globe. THese are true UTM and MGRS coordinates associated with each latitude and longitude.

My task is to regenerate these same UTM and MGRS coordinates by ignoring the facts of Datums, Shereoids, geodesics, and other projections.

I am going to assume that the earth is a perfect Spheroid.

* Equator Circumference: 40008000 meters
* Polar Circumference:   40075000 meters

```{r dataset}
rlng <- (sample.int(36000,size=500,replace=TRUE)-1)/100-180
rlat <- (sample.int(16000,size=500,replace=TRUE)-1)/100-80
samplell <- data.frame(cbind(rlng,rlat))

samplell$UTME <- NA
samplell$UTMN <- NA
samplell$z <- NA
samplell$h <- NA
samplell$mgrs <- NA

for (i in 1:nrow(samplell)){
  samplell$mgrs[i] <- latlng_to_mgrs(samplell[i,2],samplell[i,1])
  samplell$UTME[i] <- as.matrix(mgrs_to_utm(samplell[i,7]))[4]
  samplell$UTMN[i] <- as.matrix(mgrs_to_utm(samplell[i,7]))[5]
  samplell$z[i] <- as.matrix(mgrs_to_utm(samplell[i,7]))[2]
  samplell$h[i] <- as.matrix(mgrs_to_utm(samplell[i,7]))[3]
}

datatable(samplell)
```

## Square Peg; Round Hole

### Facts about UTM and MGRS

The two coordinate systems are very similar in that they are trying to overlay a rectangular coordinate system onto a sphere.

1. sldkfsldkfj
2. sdsssssssss
  + aaaaa
  + bbbbb

## Easting

### Converting a Longitude into an Easting

```{r easting}
# first go at predicting Easting from Longitude
# with dmod which is a deviation in longitude 
# from the center of the gridzone (-3 to 3 degress)

samplell <- samplell %>% mutate(latRad = cos(rlatr)*40008000,
                                dmod = rlng%%6-3,
                                e1 = dmod*latRad/360+500000,
                                e_err1 = e1 - as.numeric(UTME))

# samplell <- samplell[-c(253,436),] removed weird extreme data

samplell %>%
  ggplot() +
    geom_point(aes(rlat,e_err1,color = dmod)) +
    geom_point(aes(rlat,(-150*sin(rlat*pi/140)*sin(rlat*pi/70)-600)), 
               color = "orange")

samplell <- samplell %>% mutate(
  e2 = e1 - dmod/3*(-150*sin(rlat*pi/140)*sin(rlat*pi/70)-600),
                                e_err2 = e2 - as.numeric(UTME))

samplell %>%
  ggplot() +
    geom_point(aes(rlat,e_err2,color = dmod)) +
    geom_point(aes(rlat,(100*sin(rlat*pi/140)*sin(rlat*pi/70)-100)), 
               color = "orange")

samplell <- samplell %>% mutate(
  e3 = e2 - dmod/3*(35*cos(rlat*pi/42)/cos(rlat*pi/168)+100),
  e_err3 = e3 - as.numeric(UTME))

samplell %>%
  ggplot() +
    geom_point(aes(rlat,e_err3,color = dmod)) +
    geom_point(aes(rlat,0), 
               color = "orange")

sd(samplell$e_err3)
```



## Northing

### Converting a Latitude into a Northing

```{r northing}

# first go at predicting Northing from Latitude
# with dmod which is a deviation in longitude 
# from the center of the gridzone (-3 to 3 degress)

samplell <- samplell %>% mutate(n1 = rlat*40075000/360 +
                                  ifelse(h =="N",0,10000000),
                                n_err1 = n1 - as.numeric(UTMN))

# First try at Northing Error

a<-28000
p<-118
samplell %>%
  ggplot() +
    geom_point(aes(rlat,n_err1,color = dmod)) +
    geom_point(aes(rlat,(a*sin(rlat*pi/p))), color = "orange")

samplell <- samplell %>% mutate(
  n2 = n1 - (a*sin(rlat*pi/p)),
  n_err2 = n2 - as.numeric(UTMN))

samplell %>%
  ggplot() +
    geom_point(aes(rlat,n_err2,color = dmod)) +
    geom_point(aes(rlat,(1000*sin(rlat*pi/55))), color = "orange")


```

## WGS84

https://en.wikipedia.org/wiki/Universal_Transverse_Mercator_coordinate_system

```{r wgs84}
# αβγδεζηθικλμνξοπρςστυφχψω k0

E0 <- 500000
k0 <- .9996
f <- 1/298.257223563
n <- f/(2-f)
a <- 6378137
A <- a/(1-n)*(n^2/4)/(1-n/2)


ll2UTM2MGRS <- function(lng,lat){
  
}

lng <- 50
lat <- -40
N0 <- 0
if (lat < 0) {N0 <- 10000000}
t <- sinh(atanh(sin(lat))-2*sqrt(n)/(1+n)*atanh(2*sqrt(n)/(1+n)*sin(lat)))
nt <- atanh(sin(lng)/sqrt(1-t))
Xi <- atan(t/cos(lng))
myFun <- function(i) return(expk(i) * arfit$ar[i])
tmp2 <- sum(vapply(seq_along(arfit$ar), myFun, complex(1)))
easting = E0+k0*A*(nt+1)

```

## WGS84

```{r LL2UTM}

lng <- 45.72
lat <- -3.96
lngr <- lng*pi/180 # Longitude in radians
latr <- lat*pi/180 # Latitude in radians

# Lets go find Easting

sin1 <- pi/(180*3600) #One Second
a <- 6378137 # Equitorial Radius
b <- 6356752.3142 #Polar Radius
k0 <- .9996 #Scalar Factor Constant
zone <- floor(1/6*lng)+31 #Longitude Zone
Czone <- 6*zone - 183 #Longitude of the center of the zone
dlng <- lng - Czone # Longitude from the center of the zone
p <- dlng*3600/10000 #Hecta seconds?
e <- sqrt(1-(b/a)^2) #eccentricity
e1sq <- e^2/(1-e^2)
nu <- a/sqrt(1-(e*sin(latr))^2) #r curv 2
Kiv <- nu*cos(latr)*sin1*k0*10000 #Coef for UTM 4
Kv <- (sin1*cos(latr))^3*(nu/6)*
  (1-tan(latr)^2+e1sq*cos(latr)^2)*k0*10^12 #Coef for UTM 5
Easting <- 500000+Kiv*p+Kv*p^3

# Now let's go find Northing
n <- (a-b)/(a+b)
A0 <- a*(1-n+(5*n^2/4)*(1-n)+(81*n^4/64)*(1-n)) # Meridional Arc Length
B0 <- (3*a*n/2)*(1-n-(7*n^2/8)*(1-n)+55*n^4/64) # Meridional Arc Length
C0 <- (15*a*n^2/16)*(1-n+(3*n^2/4)*(1-n)) # Meridional Arc Length
D0 <- (35*a*n^3/48)*(1-n+11*n^2/16) # Meridional Arc Length
E0 <- (315*a*n^4/51)*(1-n) # Meridional Arc Length
S <- A0*latr - B0*sin(2*latr) + C0*sin(4*latr) - D0*sin(6*latr) + E0*sin(8*latr) # Meridional Arc
Ki <- S*k0 #Coef for UTM 1
Kii <- nu*sin(latr)*cos(latr)*sin1^2*k0*100000000/2 #Coef for UTM 2
Kiii <- ((sin1^4*nu*sin(latr)*cos(latr)^3)/24)*(5-tan(latr)^2+9*e1sq*cos(latr)^2*cos(latr)^4)*k0*10^16 #Coef for UTM 2
Northing <- Ki + Kii * p^2 + Kiii * p^4 
if (lat < 0) {Northing <- 10000000 + Northing} # In the Southern Hemisphere is Northing is measured from the south pole instead of from the equator

Easting <- round(Easting,0)
Northing <- round(Northing,0)


## Now let's turn UTM into MGRS

#Latitude gridzone letters
lgzl <- c("C","D","E","F","G","H","J","K","L","M","N","P","Q","R","S","T","U","V","W","X")

# 100km grid Square easting letter
esqls <- as.matrix(rbind(c("A","B","C","D","E","F","G","H"),
                         c("J","K","L","M","N","P","Q","R"),
                         c("S","T","U","V","W","X","Y","Z")))
colnames(esqls) <- 1:8
rownames(esqls) <- c(1,2,0)

nsqls <- c("B","C","D","E","F","G","H","J","K","L","M","N","P","Q","R","S","T","U","V","A","B","C","D","E","F","G","H","J","K","L","M","N","P","Q","R","S","T","U","V","A","B","C","D","E","F","G","H","J","K","L","M","N","P","Q","R","S","T","U","V","A","B","C","D","E","F","G","H","J","K","L","M","N","P","Q","R","S","T","U","V","A","B","C","D","E","F","G","H","J","K","L","M","N","P","Q","R","S","T","U","V")

lzc <- (lat+80-(lat+80)%%8)/8+1 # LAtitude letter zone count
lz <- lgzl[lzc]
esq <- plyr::round_any(Easting,100000,floor)/100000 #Grab the first digit off of easting
emod <- mod(zone,3) # Zone mod 3 for 100km lookup
if (emod == 0) {emod <- 3} #turn mod zero into row 3
esql <- esqls[emod,esq] #Here is the 100km easting grid letter
nsq <-  plyr::round_any(Northing,100000,floor)/100000 #Grab off the first two digits off the Northing
  if((zone %% 2) == 0) {nsq <- nsq +5} #Even grid zones increase the letter by 5
  if(nsq>99){nsq <- nsq - 20} #If it goes over the top then cycle back to "A"
nsql <- nsqls[nsq]
east <- plyr::round_any(Easting,1) - plyr::round_any(Easting,100000,floor) #Snatch the right 5 digits off the easting
easting5 <- f_pad_zero(round(east,0), width = 5, pad.char = "0") #fill the front with zeros if needed
north <- plyr::round_any(Northing,1) - plyr::round_any(Northing,100000,floor) #Snatch the right 5 digits off the northing
northing5 <- f_pad_zero(north, width = 5, pad.char = "0") #fill the front with zeros if needed

utmz <- f_pad_zero(zone, width = 2, pad.char = "0") #pad the zone with a zero, if needed
MGRS <- paste(utmz,lz,esql,nsql,easting5,northing5,sep = "") #Paste all six pieces together
hem <- "N"
if (lat < 0) {hem <- "S"}

check <- latlng_to_mgrs(lat,lng)
  
final <- as.data.frame(rbind(c(lng,lat, Easting,Northing,utmz,hem,MGRS,check))) 
colnames(final) <- c("Long","Lat","Easting","Northing","zone","Hemisphere","MGRS","Check")              

final

```

## Function LL2UTM2MGRS

```{r ll2UTM2MGRS}
pacman::p_load(tidyverse, numform)

#Function
LL2UTM2mgrs<-function(lng,lat){
  lngr <- lng*pi/180 # Longitude in radians
  latr <- lat*pi/180 # Latitude in radians
  
  # Lets go find Easting
  
  sin1 <- pi/(180*3600) #One Second
  a <- 6378137 # Equitorial Radius
  b <- 6356752.3142 #Polar Radius
  k0 <- .9996 #Scalar Factor Constant
  zone <- floor(1/6*lng)+31 #Longitude Zone
  Czone <- 6*zone - 183 #Longitude of the center of the zone
  dlng <- lng - Czone # Longitude from the center of the zone
  p <- dlng*3600/10000 #Hecta seconds?
  e <- sqrt(1-(b/a)^2) #eccentricity
  e1sq <- e^2/(1-e^2)
  nu <- a/sqrt(1-(e*sin(latr))^2) #r curv 2
  Kiv <- nu*cos(latr)*sin1*k0*10000 #Coef for UTM 4
  Kv <- (sin1*cos(latr))^3*(nu/6)*
    (1-tan(latr)^2+e1sq*cos(latr)^2)*k0*10^12 #Coef for UTM 5
  Easting <- 500000+Kiv*p+Kv*p^3
  
  # Now let's go find Northing
  n <- (a-b)/(a+b)
  A0 <- a*(1-n+(5*n^2/4)*(1-n)+(81*n^4/64)*(1-n)) # Meridional Arc Length
  B0 <- (3*a*n/2)*(1-n-(7*n^2/8)*(1-n)+55*n^4/64) # Meridional Arc Length
  C0 <- (15*a*n^2/16)*(1-n+(3*n^2/4)*(1-n)) # Meridional Arc Length
  D0 <- (35*a*n^3/48)*(1-n+11*n^2/16) # Meridional Arc Length
  E0 <- (315*a*n^4/51)*(1-n) # Meridional Arc Length
  S <- A0*latr - B0*sin(2*latr) + C0*sin(4*latr) - D0*sin(6*latr) + E0*sin(8*latr) # Meridional Arc
  Ki <- S*k0 #Coef for UTM 1
  Kii <- nu*sin(latr)*cos(latr)*sin1^2*k0*100000000/2 #Coef for UTM 2
  Kiii <- ((sin1^4*nu*sin(latr)*cos(latr)^3)/24)*(5-tan(latr)^2+9*e1sq*cos(latr)^2*cos(latr)^4)*k0*10^16 #Coef for UTM 2
  Northing <- Ki + Kii * p^2 + Kiii * p^4 
  if (lat < 0) {Northing <- 10000000 + Northing} # In the Southern Hemisphere is Northing is measured from the south pole instead of from the equator
  
  Easting <- round(Easting,0)
  Northing <- round(Northing,0)

  ## Now let's turn UTM into MGRS
  
  #Latitude gridzone letters
  lgzl <- c("C","D","E","F","G","H","J","K","L","M","N","P","Q","R","S","T","U","V","W","X")
  
  # 100km grid Square easting letter
  esqls <- as.matrix(rbind(c("A","B","C","D","E","F","G","H"),
                           c("J","K","L","M","N","P","Q","R"),
                           c("S","T","U","V","W","X","Y","Z")))
  colnames(esqls) <- 1:8
  rownames(esqls) <- c(1,2,0)
  
  nsqls <- c("B","C","D","E","F","G","H","J","K","L","M","N","P","Q","R","S","T","U","V","A","B","C","D","E","F","G","H","J","K","L","M","N","P","Q","R","S","T","U","V","A","B","C","D","E","F","G","H","J","K","L","M","N","P","Q","R","S","T","U","V","A","B","C","D","E","F","G","H","J","K","L","M","N","P","Q","R","S","T","U","V","A","B","C","D","E","F","G","H","J","K","L","M","N","P","Q","R","S","T","U","V")
  
  lzc <- (lat+80-(lat+80)%%8)/8+1 # LAtitude letter zone count
  lz <- lgzl[lzc]
  esq <- plyr::round_any(Easting,100000,floor)/100000 #Grab the first digit off of easting
  emod <- mod(zone,3) # Zone mod 3 for 100km lookup
  if (emod == 0) {emod <- 3} #turn mod zero into row 3
  esql <- esqls[emod,esq] #Here is the 100km easting grid letter
  nsq <-  plyr::round_any(Northing,100000,floor)/100000 #Grab off the first two digits off the Northing
  if((zone %% 2) == 0) {nsq <- nsq +5} #Even grid zones increase the letter by 5
  if(nsq>99){nsq <- nsq - 20} #If it goes over the top then cycle back to "A"
  nsql <- nsqls[nsq]
  east <- plyr::round_any(Easting,1) - plyr::round_any(Easting,100000,floor) #Snatch the right 5 digits off the easting
  easting5 <- f_pad_zero(round(east,0), width = 5, pad.char = "0") #fill the front with zeros if needed
  north <- plyr::round_any(Northing,1) - plyr::round_any(Northing,100000,floor) #Snatch the right 5 digits off the northing
  northing5 <- f_pad_zero(north, width = 5, pad.char = "0") #fill the front with zeros if needed
  utmz <- f_pad_zero(zone, width = 2, pad.char = "0") #pad the zone with a zero, if needed
  MGRS <- paste(utmz,lz,esql,nsql,easting5,northing5,sep = "") #Paste all six pieces together
  
  
  hem <- "N"
  if (lat < 0) {hem <- "S"}

  final <- as.data.frame(rbind(c(lng,lat, Easting,Northing,utmz,hem,MGRS))) 
  colnames(final) <- c("Long","Lat","Easting","Northing","zone","Hemisphere","MGRS")  
  
   return(final)
}

## Here is the function that turns

LL2UTM2mgrs(-138.23,-34.55)

longv <- (sample.int(36000,size=200,replace=TRUE)-1)/100-180
latv <- (sample.int(16000,size=200,replace=TRUE)-1)/100-80
samplell <- data.frame(cbind(longv,latv))


samplell$UTME <- NA
samplell$UTMN <- NA
samplell$gz <- NA
samplell$hem <- NA
samplell$mgrs <- NA

for (i in 1:nrow(samplell)){
  samplell$UTME[i] <- as.matrix(LL2UTM2mgrs(samplell$longv[i],samplell$latv[i]))[3]
  samplell$UTMN[i] <- as.matrix(LL2UTM2mgrs(samplell$longv[i],samplell$latv[i]))[4]
  samplell$gz[i] <- as.matrix(LL2UTM2mgrs(samplell$longv[i],samplell$latv[i]))[5]
  samplell$hem[i] <- as.matrix(LL2UTM2mgrs(samplell$longv[i],samplell$latv[i]))[6]
  samplell$mgrs[i] <- as.matrix(LL2UTM2mgrs(samplell$longv[i],samplell$latv[i]))[7]
}

```

## Conclusion